---
title: "Correlation and Regression"
author: "Chirantan Ganguly"
date: "27/04/2020"
output: html_document
---
## Scatterplots
Let's have a look at the relationship between height and weight through a scatterplot, using the R function plot().
```{r}
plot(women$weight,women$height,main="Heights and Weights",xlab="Weight",ylab="Height")
```

## Calculating Correlation
We can calculate the correlation in R using the function cor(), which takes your two variables as it's first argument. 
```{r}
cor(women$height,women$weight)
```

## Finding the Line
While Finding the line we use sum of the squared error we can do the same as follows:-
```{r}
y1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
y2 <- c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11)
y <- c(3, 2, 1, 4, 5, 10, 8, 7, 6, 9)
sum((y1-y)^2)
sum((y2-y)^2)
```

## Finding The Regression Coefficients in R
We can find the regression coefficients for our data using the lm() function
```{r}
money <- c(1,2,3,4,5,6,7,8,9,10)
prosocial <- c(3, 2, 1, 4, 5, 10, 8, 7, 6,9)
lm(prosocial~money)
```

## Using lm() To Add A Regression Line To Your Plot
```{r}
plot(money, prosocial, xlab = "Money", ylab = "Prosocial Behavior")
```
 After you have created your scatterplot, you can add a line using the function abline()
```{r}
plot(money, prosocial, xlab = "Money", ylab = "Prosocial Behavior")
line <- lm(prosocial ~ money)
abline(line)
```
 
## Adding a Line
We can use abline() to add any line we like, as long as the first argument is the intercept and the second is the slope.  
```{r}
plot(money, prosocial, xlab = "Money", ylab = "Prosocial Behavior")
abline(mean(prosocial),0)
abline(line)
```

## Calculating r^2
The reduction in error by using the regression line compared to the mean line tells us how well the independent variable (money) predicts the dependent variable (prosocial behaviour). This reduction is predicted by r^2 : -
```{r}
r<-(cor(money,prosocial))^2
r
```


## Putting it all Togather: Regression
Lets try to put all what we have learned together:-

```{r}
money <- c(4, 3, 2, 2, 8, 1, 1, 2, 3, 4, 5, 6, 7, 9, 9, 8, 12)
education <- c(3, 4, 6, 9, 3, 3, 1, 2, 1, 4, 5, 7, 10, 8, 7, 6, 9)
r<-cor(money,education)
r
line<-lm(money~education)
line
plot(education,money,main="My Scatterplot")
abline(line)
```

