---
title: "Introduction to R continued"
author: "Chirantan Ganguly"
date: "26/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R continued
___

## Functions: -

### Getting Help

While working with R if we need some additional info about what a function does, we can get detailed info about it in the following ways.
```{r}
  help(mean)  #asking for info about the function mean
  ?mean       #another way of doing the same thing
```

We can also know what arguements a function takes in the following way:-
```{r}
  args(mean)
```

### More about FUnctions

Now that we have seen a bit about the function mean(). Lets actually calculate mean()
We can calculate mean by 2 methods:- 
1. Matching by name: - mean(x= data)
```{r}
  grades<-c(8.5, 7, 9, 5.5, 6)
  mean(x= grades)
```
2. Matching by position : - mean(data)
```{r}
mean(grades)
```
If our data set has missing values then mean() will throw an error if not specified as follows: -
```{r}
grades <- c(8.5, 7, 9, NA, 6)
mean(grades)
mean(x=grades, trim = 0, na.rm = TRUE)
```

### Making Your own Functions

Apart from builtin functions, we can make functions of our own like: -
```{r}
multiply_a_b <- function(a,b){
  return(a*b)
}
multiply_a_b(3,7)
```


## Getting your data into R

One important thing before you actually do analyses on your data, is that you will need to get your data into R. R contains many functions to read in data from different formats. Some Examples are as follows: -

```{r}
cars<-read.csv("http://s3.amazonaws.com/assets.datacamp.com/course/uva/mtcars.csv")
head(cars)
```
Sometimes the sperator is not what we expect it to be so we can easily change the separator by: -
```{r}
cars<-read.csv("http://s3.amazonaws.com/assets.datacamp.com/course/uva/mtcars_semicolon.csv", sep = ';')
head(cars)
```

## Working directories in R

If you would work with R studio on your own computer, you would probably like to read in local files.

\- getwd(): This function will retrieve the current working directory for the user
\- setwd(): This functions allows the user to set her own working directory

```{r}
getwd()
setwd("F:/Education/College/Computer Science/Data_Science")
```

## Checking files in your working directory

```{r}
list.files()
cars<-read.csv("mtcars_semicolon.csv", sep=';')
head(cars)
```

## Importing R packages

```{r}
library(ggplot2)
require(ggplot2)
```

