---
title: "The Basics of R"
author: "Chirantan Ganguly"
date: "26/04/2020"
output: html_document
---
# The Basics Of R (Data Camp)
I have tried to incorporate a summary of The Basics of R in this file for future reference and practice.

## Little Arithmetics with R

```{r}
  3+4
  5-5
  3*5
  (5+5)/2
  2^5
  28%%6
```

## Basic Data Types

```{r}
  my_numeric<- 42
  class(my_numeric)
  my_character<-"forty two"
  class(my_character)
  my_logical<-is.character(my_character)
  my_logical
  class(my_logical)
```

## Coercion

```{r}
  var1<-TRUE
  var2<-3
  var3<-"15"
  var1_char<-as.character((var1))
  var1_char
  var2_log<-as.logical((var2))
  var2_log
  var3_num<-as.numeric(var3)
  var3_num
```

## Vectors

```{r}
  numeric_vector<-c(1,2.3,4.2,5,10,49,67)
  numeric_vector
  character_vector<-c("Chirantan","Ganguly","Calcutta")
  character_vector
  log_vector<-c(TRUE, FALSE, FALSE, TRUE)
  log_vector
```

**Accessing Vector Elements**

```{r}
  numeric_vector[3]
  character_vector[c(1,3)]
  log_vector[c(1:3)]
```

## Selection by Comparison

```{r}
  larger_than_ten<-numeric_vector>10
  larger_than_ten
  numeric_vector[larger_than_ten]
```

## Matrices

```{r}
  m<-matrix(1:20, byrow=TRUE, nrow=5, ncol=4)
  m
```

## Factor

**Used to fix the number of possible values a vector can take**
```{r}
  student_status<-c("student","not student", "not student", "student", "student", "student")
  categorical_student<-factor(student_status)
  categorical_student
```

## DataFrames

```{r}
mtcars
```

**Inspecting DataFrames**

```{r}
  head(mtcars)
  tail(mtcars)
  str(mtcars)
  dim(mtcars)
  colnames(mtcars)
```

## Contructing Datafame od your own

```{r}
planets <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)  #relative to earth
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)   #relative to earth
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
planet_df<-data.frame(planets, type, diameter, rotation, rings)
planet_df
```

## Indexing and Selecting Columns from dataframe

```{r}
planet_df[1,c(2,3)]
planet_df$diameter
```

## Lists

```{r}
my_vector <- 1:10 
my_matrix <- matrix(1:9, ncol = 3)
my_df <- mtcars[1:10,]
my_list<-list(my_vector, my_matrix, my_df)
my_list
```

## Selecting Elements from a List

```{r}
my_list[2]
my_list[[3]][,1]
```

